---
title: "AML_Drug_Comnbination"
output: 
  pdf_document:
    fig_caption: yes
fontsize: 11pt
geometry: margin=1in
graphics: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE, fig.pos= "h")
```

# Library   
```{r libraries and reading files, include=FALSE}
library(kableExtra)
options(knitr.table.format = "latex")
library(readr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(tidyr)
library(fpc)
library(grid)
library(reshape2)
library(topGO)
library(dplyr)
library(visNetwork)
library(pheatmap)
library(Rtsne)
library(gridExtra)
library(fpc)
library(fgsea)
library(data.table)
library(reactome.db)
library(org.Hs.eg.db)
library(ggnewscale)
library(NIMAA)
library(viridis)
library(viridisLite)
library(ggpubr)
library(rstatix)
library(synergyfinder)
library(readxl)
library(reader)
```

# Data reading, cleaning, preprocessing

**DSRT results** data set includes 967,921 experiments. Here, each row of data set was considered as an experiment. In this study we just consider the results of experiments on AML sample patients that inhibition measured by "Dead cells(CellTox Green) readout". The experiments with empty or NAN values on the `drug` and `disease.stage` fields are removed. The `remission`  samples and `dubious` experiments were removed as well. The experiments with empty `add.ons` values are considered. We analysed only `BM` samples. Here, we have three kind of mediums `CM 0.125`, `CM 0.25`,and `MCM`. We removed experiments on `MCM` medium. After data cleaning we get `aml_dead` with 75,278 rows and `healthy_dead` with 5,911 rows.

## Dose-response Preprocessing

For each experiment the vector of dose-response was normalized. At first values larger than 100 set to 100 and values below 0 set to zero. In order to keep decreasing or increasing pattern of response function, the middle values are smoothed.  

```{r normalized_dose-response}
aml_raw <- readRDS(file = "data/rds/aml_raw.rds");
healthy_raw <- readRDS(file = "data/rds/healthy_raw.rds");
smooth_response <- function(dose_response,readout) {options(digits=2)
  A <- as.numeric(unlist(strsplit(as.character(dose_response),";")))
  A[A > 100] <- 100;A[A < 0] <- 0
  An <- A[-5];Ab <- append(0, An);diff <- A - Ab;
  idx <- which(diff == -100);idxm <-idx[!(idx == 1 | idx ==5)];
  idx5 <- idx[(idx ==5)]
  A[idxm-1] <- 0
  An <- A[-5];Ab <- append(0, An);diff <- A - Ab;
  idx <- which(diff < 0);idxm <-idx[!(idx == 1 | idx ==5)];
  idx5 <- idx[(idx ==5)]
  A[idxm] <- (A[idxm-1] + A[idxm+1]) / 2; A[idx5] <- A[4]
  A <- round(A,digits = 2)
  vectStr=paste(as.character(A), sep="' '", collapse=";")
  df <- data.frame(vectStr)
  colnames(df) <- c("response")
  return(df)
}

norm_response <- data.frame(do.call(rbind,lapply(1:dim(aml_raw)[1], function(x) smooth_response (aml_raw$dose.response[x],aml_raw$readout[x]))))
aml <- cbind(aml_raw,norm_response)
rm(norm_response)
#saveRDS(aml, file = "data/rds/aml.rds")  
norm_response <- data.frame(do.call(rbind,lapply(1:dim(healthy_raw)[1], function(x) smooth_response (healthy_raw$dose.response[x],healthy_raw$readout[x]))))
healthy <- cbind(healthy_raw,norm_response)
rm(norm_response)
#saveRDS(healthy, file = "data/rds/healthy.rds") 
```

# Full submatrix

## Mean value of dose response calculation
The mean value of drug response for each experiment is calculated. Then the matrix of sample-by-drug in which its entries are the mean value of the effect of drug on specific sample is constructed. This matrix includes 199 samples-by 625 drugs.
```{r mean_value}
aml <- readRDS(file = "data/rds/aml.rds");
healthy <- readRDS(file = "data/rds/healthy.rds")
#-----------------------calculation of mean value of drug response--------------------------------
splitdat = do.call(rbind, strsplit(aml$response, ";"))
splitresp = data.frame(apply(splitdat, 2, as.numeric))
names(splitresp) = paste("dose", 1:5, sep = "")
aml <- cbind(aml,splitresp)
sample_drug_response <- aml %>%
  group_by(sample, drug) %>% 
  dplyr::summarise(across(dose1:dose5, mean)) # mean on replicates
sample_drug_response <- cbind(sample_drug_response,mean=rowMeans(sample_drug_response[3:7], na.rm=TRUE))
#saveRDS(sample_drug_response, file = "data/rds/sample_drug_response.rds") 
#------------------------selecting sample, drug, and mean value and spread it-----------------------
sample_drug_mean <- sample_drug_response %>% dplyr::select(sample,drug, mean) %>% spread(drug, mean)
sample_drug_matrix_mean <- (as.matrix(sample_drug_mean[,-c(1)]))
rownames(sample_drug_matrix_mean) <- sample_drug_mean$sample
#saveRDS(sample_drug_matrix_mean,file ="data/rds/sample_drug_matrix_mean.rds")
```

## Full AML submatrix extraction

The full sub matrix with no missing entries is extracted and then min_max normalization procedure implemented on it. The matrix includes 81 samples and 296 drugs.
```{r Full_submatrix}
sample_drug_matrix_mean <- readRDS(file = "data/rds/sample_drug_matrix_mean.rds")
healthy <- readRDS(file = "data/rds/healthy.rds")
#-----------------------extraction of full submatrix--------------------------------------------
sub_matrices <- extractSubMatrix(sample_drug_matrix_mean,
                                 shape = "All",
                                 verbose = FALSE,
                                 palette = "Greys",
                                 row.vars = NULL,
                                 col.vars = NULL, 
                                 bar = 1,
                                 plot_weight = FALSE,
                                 print_skim = F)

max_rec_mean <- sub_matrices$Rectangular_element_max
#saveRDS(max_rec_mean, file = "data/rds/max_rec_mean.rds")
drugs <- colnames(max_rec_mean)
healthy_rec <- healthy[healthy$drug %in% drugs,]
#saveRDS(healthy_rec, file = "data/rds/healthy_rec.rds") 
#-------------------minmax normalization-----------------------------
max_rec_minmax <- (max_rec_mean - min(max_rec_mean))/(max(max_rec_mean)-min(max_rec_mean))
#saveRDS(max_rec_minmax, file = "data/rds/max_rec_minmax.rds")
rm(list=c("aml","splitdat","splitresp","sample_drug_matrix","drugs"))
```

## Healthy induced submatrix

The results of drug response on healthy samples trimmed according to 296 drugs in AML full submatrix. The mean value of drug response for each experiment is calculated. This matrix includes 7 samples-by 296 drugs.
```{r}
healthy_rec <- readRDS(file = "data/rds/healthy_rec.rds")
#-----------------------calculation of mean value of drug response--------------------------------
splitdat = do.call(rbind, strsplit(healthy_rec$response, ";"))
splitresp = data.frame(apply(splitdat, 2, as.numeric))
names(splitresp) = paste("dose", 1:5, sep = "")
healthy_rec <- cbind(healthy_rec,splitresp)
sample_drug_response_healthy <- healthy_rec %>%
  group_by(sample, drug) %>% 
  dplyr::summarise(across(dose1:dose5, mean))# mean on biological replicates
#---------------------mean on dose response values--------------------
sample_drug_response_healthy <- 
  cbind(sample_drug_response_healthy,mean=rowMeans(sample_drug_response_healthy[3:7], na.rm=TRUE))
#----------------------minmax normalization-------------------------
sample_drug_response_healthy <- sample_drug_response_healthy  %>%
  transform(minmax = (mean - min(.$mean)) / (max(.$mean) - min(.$mean)))
#saveRDS(sample_drug_response_healthy, file = "data/rds/sample_drug_response_healthy.rds") 
```

# Network analysis
## Bipartite Network construction

The bipartite network is constructed using full sub_matrix. The data matrix A = sample_drug_wmean was used to construct a weighted bipartite network where V1 = S was set of 81 samples, and V2 = D consisted of 296 drugs. The weight of the edge that joins node si (sample i) and node dj (drug j) was the ij entry of the matrix A. A weighted bipartite network was built, with two parts: samples and compounds, and weight representing the inhibition rate.
```{r}
max_rec <- readRDS(file = "data/rds/max_rec_minmax.rds")
# ---------this dataframe includes the final minmax normalized aml sample drugs-----
sample_drug_wmean <- melt(as.matrix(max_rec))
colnames(sample_drug_wmean) <- c("sample","drug","w_mean")
#------------network construction based on sample_drug_wmean dataframe-----
bipart_wmean <-  sample_drug_wmean %>% graph_from_data_frame(directed = FALSE)
V(bipart_wmean)$type[1:81] <- FALSE
V(bipart_wmean)$type[82:377] <- TRUE
E(bipart_wmean)$weight <- sample_drug_wmean[,3]
V(bipart_wmean)$color <- c("steel blue", "orange")[V(bipart_wmean)$type+1]
V(bipart_wmean)$shape <- c("square", "circle")[V(bipart_wmean)$type+1]
#saveRDS(bipart_wmean, file = "data/rds/bipart_wmean.rds")
#saveRDS(sample_drug_wmean, file = "data/rds/sample_drug_wmean.rds")
```

### Visualization

Bipartite network visualization
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot(sample_drug_net_median_sq, vertex.label=NA, vertex.size=(2-V(sample_drug_net_median_sq)$type)*4, layout = layout_nicely)
library(GGally )
library(cowplot )
library(ggpubr)
bipart_wmean <- readRDS(file = "data/rds/bipart_wmean.rds")
sample_drug_wmean <- readRDS(file = "data/rds/sample_drug_wmean.rds")
bg <- ggnet2(bipart_wmean, size = 2.5, color = (4 - V(bipart_wmean)$type), edge.alpha = 0.02, palette = "Set1", 
layout.par = list(cell.jitter = .08) )+
#  theme(legend.position = "none")+
#  scale_color_discrete(name = "Node", labels = c("Sample", "Drug"))
scale_color_manual(name = "", labels = c("Patient-derived Sample", "Chemical Compound"), values=c("cornflowerblue", "darkorange"))
wg <- ggplot(sample_drug_wmean, aes(1, w_mean))+
  geom_jitter( shape = 21, size=2, color="darkorange",alpha=0.1)+
     geom_violin(alpha = 0.4, size=1, color="cornflowerblue", draw_quantiles = c(0.25, 0.5, 0.75)) +
     theme_pubr()+ ylim(0,0.05)+ coord_flip() + xlim(0.5, 4)+ 
   ylab("")+
  annotate("text", label = 'atop(bold("Weight of bigraph edges"))', 
           x = 1.4, y = 0.025, size = 6, colour = "black", parse = TRUE)+
  theme(axis.line.y = element_blank(), axis.ticks.y = element_blank(), 
        axis.text.y = element_blank(), axis.title.y = element_blank(),
        axis.line.x = element_blank(), axis.title.x = element_text(face = "bold"),
        plot.title = element_text(face = "bold", size = 24))
  #geom_violin()+ ylim(0,0.05)+ coord_flip()+theme()
wg

inset_plot <- ggdraw()+
  draw_plot(wg)+
  draw_plot(bg, x = 0.1, y = 0.35, width = 0.8, height = 0.6)
inset_plot
# forGephi <- as_edgelist(sample_drug_net_median_sq)
# colnames(forGephi) <- c("Source", "Target")
# write.csv(forGephi, file = "gephi2.csv", row.names = FALSE, col.names = FALSE, na = "", sep = ",")
#saved in pdf format in 9.5X10.5 inch as bipartite.pdf

```

## Projection of Bipartite network on drugs

A bipartite network can be projected into two different types of unipartite networks containing nodes of only one type. The projection of the bipartite network onto the "drug" node set was calculated: "proj_drug_network"
```{r}
sample_drug_wmean <- readRDS(file = "data/rds/sample_drug_wmean.rds")
sample_drug_net_mean_incd <- sample_drug_wmean %>% spread(key = drug, value = w_mean) %>% column_to_rownames("sample") %>% as.matrix
dim(sample_drug_net_mean_incd)
#str(sample_drug_net_mean_incd)
proj_drug <- t(sample_drug_net_mean_incd) %*% (sample_drug_net_mean_incd)
#dim(proj_drug)
proj_drug_network <- graph_from_adjacency_matrix((proj_drug), mode = "upper", weighted = TRUE, diag = FALSE)
#igraph::components(proj_drug_network)
#saveRDS(proj_drug_network, file = "data/rds/proj_drug_network.rds")
drug_proj_df <- as_data_frame(proj_drug_network)
#saveRDS(drug_proj_df, file = "data/rds/drug_proj_df.rds")
#write_csv(data.frame(proj_drug_med_sq), "Fig4.csv")
```

### Removing weak edges

 Only edges with a weight greater than the median of similarities were kept in order to consider them strong enough edges in the projected network.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
proj_drug_network <- readRDS(file = "data/rds/proj_drug_network.rds")
#-----------------------------remove edges < median(weights)------------------
proj_drug_network_del <- delete_edges(proj_drug_network,which(E(proj_drug_network)$weight <= 0.2))
proj_drug_network_del_w <- proj_drug_network_del
#--------------------change weighted network into unweighted network---------------
E(proj_drug_network_del)[weight > 0.2]$weight <- 1
degrees.G <- igraph::degree(g)

edge_density(proj_drug_network_del, loops = FALSE)
transitivity(proj_drug_network_del)
components(proj_drug_network_del)$no
#saveRDS(proj_drug_network_del, file = "data/rds/proj_drug_network_del.rds")
#saveRDS(proj_drug_network_del_w, file = "data/rds/proj_drug_network_del_w.rds")
```

### Drug-drug similarity Network Clustering

The projected network was clustered using louvain algorithm. We found two clusters of drugs g1 = 155 and g2 = 141 sizes.

```{r fast_greedy, echo=FALSE, message=FALSE, warning=FALSE}
proj_drug_network_del <- readRDS(file = "data/rds/proj_drug_network_del.rds")
shortpath <- distances(proj_drug_network_del,
                       v = V(proj_drug_network_del), 
                       to = V(proj_drug_network_del), 
                       mode = c("all"),
                       algorithm = c("dijkstra"))
g <- proj_drug_network_del
#---------------louvain clustering-------------------------------------------
community_lou <- cluster_louvain(g)
md <- round(modularity(g, membership(community_lou),weights = NULL),3)#modularity index
sz <- sizes(community_lou)# number of communities
V(g)$memb <- membership(community_lou)
#-----------------------------------------------------------------------------
#cl_lou <- cluster.stats(d = shortpath,
#                        membership(community_lou), 
#                       alt.clustering = NULL,
#                        noisecluster=FALSE)
#sil_lou <- cl_lou$clus.avg.silwidths;mu_sl <- round(mean(sil_lou),3)
#---------------induced subgraphs on each cluster-------------------
g1 <- induced.subgraph(g,vids=V(g)[V(g)$memb ==1])
g2 <- induced.subgraph(g,vids=V(g)[V(g)$memb ==2]) 
#length(V(g)[V(g)$memb ==1])
#length(V(g)[V(g)$memb ==2])
#saveRDS(g, file = "data/rds/g.rds")
#saveRDS(g1, file = "data/rds/g1.rds")
#saveRDS(g2, file = "data/rds/g2.rds") 
```
# Computational Validation

To accomplish intra-cluster homogeneity analysis, we employed two computational methods. The first method identified the significant difference between biological pathways of drug targets’ protein targets at each cluster, while the second evaluated drug chemical structure similarity at each cluster. 
Using the drug-target common (DTC) database we built a drug-target network, which was a bipartite network in which each link connects drugs to their protein targets. "drug_target" file includes protein targets of drugs at this study. 

##  protein target of drugs at each cluster

Protein targets of drugs at each cluster identified. "dtc_fimm_cl_list1" includes protein targets of drugs at cluster 1 and "dtc_fimm_cl_list2" includes protein targets of drugs at cluster 2.
```{r}
g <- readRDS(file = "data/rds/g.rds")
drug_target <- readRDS(file= "data/rds/drug_target.rds")
drug1 <- as.character(V(g)$name[V(g)$memb == 1])
drug2 <- as.character(V(g)$name[V(g)$memb == 2])
dtc_fimm_cl1 <- drug_target %>% filter(drug  %in%  drug1)
dtc_fimm_cl2 <- drug_target %>% filter(drug  %in%  drug2)
dtc_fimm_cl_list1 <- dtc_fimm_cl1 %>% split(list(.$drug))
dtc_fimm_cl_list2 <- dtc_fimm_cl2  %>% split(list(.$drug))
dtc_fimm_cl_list <- drug_target %>% split(list(.$drug))
drug_untargeted <-data.frame(drug = V(g)$name [!V(g)$name %in% unique(drug_target$drug)])
#saveRDS(dtc_fimm_cl_list1, file = "data/rds/dtc_fimm_cl_list1.rds")
#saveRDS(dtc_fimm_cl_list2, file = "data/rds/dtc_fimm_cl_list2.rds")
#saveRDS(dtc_fimm_cl_list, file = "data/rds/dtc_fimm_cl_list.rds")
```

## Frequency of targetted proteins by drugs at each cluster.

The frequency of protein targets by drugs at each cluster is calculated and saved at "freq_proteins" file.
```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE }
dtc_fimm_cl_list <- readRDS(file = "data/rds/dtc_fimm_cl_list.rds")
dtc_fimm_cl_list1 <- readRDS(file = "data/rds/dtc_fimm_cl_list1.rds")
dtc_fimm_cl_list2 <- readRDS(file = "data/rds/dtc_fimm_cl_list2.rds")
#union of all protein targets in cluster 1:uni1 
df_dtc_fimm <- as.data.frame(do.call(rbind, dtc_fimm_cl_list))
colnames(df_dtc_fimm) <- c("drug","protein")
df_dtc_fimm1 <- as.data.frame(do.call(rbind, dtc_fimm_cl_list1))
colnames(df_dtc_fimm1) <- c("drug","protein")
df_dtc_fimm2 <- as.data.frame(do.call(rbind, dtc_fimm_cl_list2))
colnames(df_dtc_fimm2) <- c("drug","protein")
uni_proteins <- unique(df_dtc_fimm$protein)# number of protein targets 
uni_proteins1 <- unique(df_dtc_fimm1$protein)# number of protein targets of cluster 1
uni_proteins2 <- unique(df_dtc_fimm2$protein)# number of protein targerts of cluster 2
freq_proteins <- data.frame(protein = uni_proteins,freq1 = rep(0,length(uni_proteins)),
                            freq2=rep(0,length(uni_proteins)))
for (i in seq_along(dtc_fimm_cl_list1)) {
  unii <- unique(as.vector(unique(unlist(strsplit(as.character(dtc_fimm_cl_list1[[i]]$target),",")))))
    if (!is_empty(unii)){
      if (!anyNA(unii)){
        ind <-  freq_proteins$protein %in% unii
        freq_proteins$freq1[ind] <- freq_proteins$freq1[ind] +1 
      }
    }
}
for (i in seq_along(dtc_fimm_cl_list2)) {
  unij <- unique(as.vector(unique(unlist(strsplit(as.character(dtc_fimm_cl_list2[[i]]$target),",")))))
    if (!is_empty(unij)){
      if (!anyNA(unij)){
        ind <-  freq_proteins$protein %in% unij
       freq_proteins$freq2[ind] <- freq_proteins$freq2[ind] +1 
      }
    }
}
freq_proteins$sfreq1 <- (freq_proteins$freq1)/length(dtc_fimm_cl_list1)
freq_proteins$sfreq2 <- (freq_proteins$freq2)/length (dtc_fimm_cl_list2)
freq_proteins$ratio <-  ((freq_proteins$sfreq1)/(freq_proteins$sfreq2))
#saveRDS(freq_proteins, file = "data/rds/freq_proteins.rds")
#saveRDS(df_dtc_fimm1, file = "data/rds/df_dtc_fimm1.rds")
#saveRDS(df_dtc_fimm2, file = "data/rds/df_dtc_fimm2.rds")
```

## logRatio 
You do not need to run this chunk. All variables already saved. In this chunk we extracted the name of proteins, calculated logratio, and made a named vector "protein_list22" to do GSEA analysis. All variables needed for GSEA already saved.
To better understand the protein targets of drugs in each cluster, we assign a score to each protein based on the number of distinct drugs targeting that protein in clusters 1 and 2. Let  "freq_proteins$freq1" ("freq_proteins$freq2" ) denotes the number of unique drugs in cluster C1 (C2) targeted a particular protein P. The score of protein P, defined by freq_proteins$ratio . Proteins with a score greater than log(2) are considered to be preferentially targeted by drugs in cluster 1, denoted by PPT1. Similarly, PPT2 proteins have a
score of less than log(0.50)
```{r}
freq_proteins <- readRDS(file = "data/rds/freq_proteins.rds")
###########pathway analysis using ratio of freq
freq_proteins$ratio <- log(freq_proteins$ratio)
freq_proteins$ratio[freq_proteins$freq2==0]=max(freq_proteins$ratio[is.finite(freq_proteins$ratio)])+1 
freq_proteins$ratio[freq_proteins$ratio==-Inf]= min(freq_proteins$ratio[is.finite(freq_proteins$ratio)])-1 
freq22_proteins <- dplyr::filter(freq_proteins,((freq_proteins$ratio >= log(2) ) |(freq_proteins$ratio <= log(0.5) )))
#write.csv(freq22_proteins$protein,"data/xls/proteins_22.csv", row.names = FALSE)
#saveRDS(freq22_proteins, file = "data/rds/freq22_proteins.rds")
protein_genename <- read.delim("data/xls/protein_genename.txt")
colnames(protein_genename) <-  c("protein", "genename")
protein_geneid <- read.delim("data/xls/protein_geneid.txt")
colnames(protein_geneid) <- c("protein","geneid")
ids_22 <- left_join(protein_genename,protein_geneid,by="protein")
#saveRDS(ids_22, file = "data/rds/ids_22.rds")
freq22_protein_gene <- left_join(ids_22,freq22_proteins,by= "protein")
#saveRDS(freq22_protein_gene, file = "data/rds/freq22_protein_gene.rds")
original_gene_list <- freq22_proteins[,"ratio"]
names(original_gene_list) <- (freq22_proteins[,1])
protein_list22 <- na.omit(original_gene_list)
protein_list22 = sort(protein_list22, decreasing = TRUE)# Making named variable includes the name of proteins and  associated freqs
#saveRDS(protein_list22, file = "data/rds/protein_list22.rds")
geneid_list <- freq22_protein_gene[,"ratio"]
names(geneid_list) <- (freq22_protein_gene[,3])
geneid_list22 <- na.omit(geneid_list)
geneid_list22 = sort(geneid_list22, decreasing = TRUE)# Making named variable includes the name of proteins and  associated freqs
#saveRDS(geneid_list22, file = "data/rds/geneid_list22.rds")
genename_list <- freq22_protein_gene[,"ratio"]
names(genename_list) <- (freq22_protein_gene[,2])
genename_list22 <- na.omit(genename_list)
genename_list22 = sort(genename_list22, decreasing = TRUE)# Making named variable includes the name of proteins and  associated freqs
#saveRDS(genename_list22, file = "data/rds/genename_list22.rds")
```

## KEGG12
KEGG pathway analysis using "clusterProfiler" package on "protein_list22". Do not need to run this chunk! We already saved
all files needed for kegg pathway visualization in the next chunk. 
```{r}
library(clusterProfiler)
freq22_proteins <- readRDS(file = "data/rds/freq22_proteins.rds")
protein_list22 <- readRDS(file = "data/rds/protein_list22.rds")
#library(organism, character.only = TRUE)
# keytypes(org.Hs.eg.db)
# organism = org.Hs.eg.db
# kegg <- gseKEGG(geneList = protein_list22,
#                organism = "hsa",
#                nPerm        = 10000,
#                minGSSize    = 10,
#                maxGSSize    = 500,
#                pvalueCutoff = 0.05,
#               pAdjustMethod = "none",
#                keyType = "uniprot")
#saveRDS(kegg, file = "data/rds/kegg.rds")
kegg <- readRDS(file = "data/rds/kegg.rds")

g <- dotplot(kegg, showCategory=150, font.size = 9,split=".sign")
kegg12 <- g$data

kegg_12 <- kegg12 %>% 
    mutate(protein = strsplit(as.character(kegg12$core_enrichment), "/")) %>% 
    unnest(protein) %>%
    left_join(freq22_proteins, by = "protein") %>%
    dplyr::select(c(1,2,7,13,16,17,18,21)) %>%
    mutate(freq_path = rep(1,886)) %>%
    mutate(protein_group = ifelse(ratio >= log(2) , "PPT1","PPT2"))
    #%>%  dplyr::filter(p.adjust< 0.01) 

kegg_prot_w <- kegg_12 %>% 
  group_by(Description,protein_group,.sign) %>% 
  summarise(weight = sum(freq_path))
kegg_prot_w_15 <- kegg_prot_w %>% dplyr::filter(weight > 10)
ind_filter <- kegg_prot_w$Description %in% kegg_prot_w_15$Description
kegg_protein22 <- kegg_prot_w[ind_filter,]
#saveRDS(kegg_protein22, file = "data/rds/kegg_protein22.rds")
protein_g1 <- kegg_12 %>% filter(protein_group =="PPT1")
length(unique(protein_g1$protein))
protein_g2 <- kegg_12 %>% filter(protein_group =="PPT2")
length(unique(protein_g2$protein))
```

### Visualization
Enriched KEGG pathways of proteins PPT1 and PPT2.
```{r}
library(ggalluvial)
kegg_protein22 <- readRDS("data/rds/kegg_protein22.rds")
#%%%%%%%%%%%Visualization%%%%%%%%%%%%%%%%%
#, fontface = "bold"
keg12 <- ggplot(data = kegg_protein22,
       aes(axis1 = protein_group, axis2 = Description, y = weight)) +
  geom_alluvium(aes(fill = Description), width = 5/16,alpha = 0.4, knot.pos = 0.4) +
        geom_stratum()+
    scale_x_discrete(limits = c(),expand = c(0.01, 0.4))+
  geom_text(stat = "stratum",size=5,family = "serif",
            aes(label = after_stat(stratum))) +
  theme_void()+
  theme( legend.position = "none" ) +
  ggtitle("A) KEGG enriched pathways") +
 theme(plot.title = element_text(hjust = 0.35, vjust=2.12))

```

## BP12
Biological Processes using clusterProfiler package on "protein_list22" variable. Do not need to run this chunk! We already saved
all files needed for kegg pathway visualization in the next chunk.
```{r}
protein_list22 <- readRDS(file = "data/rds/protein_list22.rds")
library(clusterProfiler)
#   keytypes(org.Hs.eg.db)
#   organism = org.Hs.eg.db
#   gseBP <- gseGO(geneList=protein_list22,
#                ont ="BP",
#                keyType = "UNIPROT",
#                nPerm = 10000,
#                minGSSize = 10,
#                maxGSSize = 500,
#                pvalueCutoff = 0.05,
#                verbose = TRUE,
#                OrgDb = organism,
#                pAdjustMethod = "fdr" )
# saveRDS(gseBP, file = "data/rds/gseBP.rds")
gseBP <- readRDS(file = "data/rds/gseBP.rds")
freq22_proteins <- readRDS(file = "data/rds/freq22_proteins.rds")
g <- dotplot(gseBP, showCategory=133, font.size = 9,split=".sign")
bp12 <- g$data

bp_12 <- bp12 %>% 
    mutate(protein = strsplit(as.character(bp12$core_enrichment), "/")) %>% 
    unnest(protein) %>%
    left_join(freq22_proteins, by = "protein") %>%
    dplyr::select(c(1,2,7,13,16,17,18,21)) %>%
    mutate(freq_path = rep(1,3034)) %>%
    mutate(protein_group = ifelse(ratio >= log(2) , "PPT1","PPT2"))
    #%>%  dplyr::filter(p.adjust< 0.01) 

bp_prot_w <- bp_12 %>% 
  group_by(Description,protein_group,.sign) %>% 
  summarise(weight = sum(freq_path)) %>% dplyr::filter(weight > 10)

# fdr_go_bp12_active <- bp12 %>% dplyr::filter(.sign =="activated") %>%  dplyr::select(ID,p.adjust)
# fdr_go_bp12_supp <- bp12 %>% dplyr::filter(.sign =="suppressed") %>%  dplyr::select(ID,p.adjust)
#  
#  write.csv(fdr_go_bp12_active,"data/xls/GO/fdrgo_bp12_active.csv", row.names = FALSE)
#  write.csv(fdr_go_bp12_supp,"data/xls/GO/fdrgo_bp12_supp.csv", row.names = FALSE)

bp_prot_w_15_g1 <- bp_prot_w[c(36,13,22,56,50,49,53,52),] %>%
  dplyr::filter(weight > 10)
bp_prot_w_15_g2 <- bp_prot_w[c(90,17,29,20,11,130,25),] 
ids <- c(bp_prot_w_15_g1$Description,bp_prot_w_15_g2$Description)
ind_filter <- bp_prot_w$Description %in% (ids)
bp_protein22 <- bp_prot_w[ind_filter,]
saveRDS(bp_protein22,file = "data/rds/bp_protein22.rds" )
```

### Visualization
Biological Processes of proteins PPT1 and PPT2.
```{r}
library(ggalluvial)
bp_protein22 <- readRDS("data/rds/bp_protein22.rds")
bp12 <- ggplot(data = bp_protein22,
       aes(axis1 = protein_group, axis2 = Description, y = weight)) +
  geom_alluvium(aes(fill = Description), width = 5/16,alpha = 0.4, knot.pos = 0.4) +
        geom_stratum()+
    scale_x_discrete(limits = c(),expand = c(0.01, 0.4))+
  geom_text(stat = "stratum",size=5,family = "serif", aes(label = after_stat(stratum))) +
  theme_void()+
  theme( legend.position = "none" ) +
  ggtitle("B) Biological Process") +
   theme(plot.title = element_text(hjust = 0.28, vjust=2.12))
# size 13 X 5 for save in pdf
```

merge two figures KEGG and biological processes.
```{r}
library("cowplot")
library("grid")

merg <- plot_grid(keg12, bp12,
          ncol = 2,
          labels = c("A","B"))
#saved 22x10 pdf file
```

## Protein target groups

We grouped proteins into 4 clusters. freqij are proteins that are targeted by at lease i drugs from cluster 1 and j drugs from cluster 2. p1 includes proteins that are mainly targeted by drugs in cluster 1 and p2 includes proteins that are mainly targeted by drugs in cluster 2. We do enrichment analysis on p1 and p2 to show that they are at different pathways.

```{r}
freq_proteins <- readRDS(file = "data/rds/freq_proteins.rds")
freq22_proteins <- dplyr::filter(freq_proteins,(freq_proteins$freq1 > 2 & freq_proteins$freq2 > 2))
#saveRDS(freq22_proteins, file = "data/rds/freq22_proteins.rds")
freq12_proteins <- dplyr::filter(freq_proteins,(freq_proteins$freq1 <= 2 & freq_proteins$freq2 > 2))
#saveRDS(freq12_proteins, file = "data/rds/freq12_proteins.rds")
freq21_proteins <- dplyr::filter(freq_proteins,(freq_proteins$freq1 > 2 & freq_proteins$freq2 <= 2))
#saveRDS(freq21_proteins, file = "data/rds/freq21_proteins.rds")
freq11_proteins <- dplyr::filter(freq_proteins,(freq_proteins$freq1 <= 2 & freq_proteins$freq2 <= 2))
#saveRDS(freq11_proteins, file = "data/rds/freq11_proteins.rds")
#write.csv(freq12_proteins$protein, "data/xls/p2.csv" ,row.names = FALSE, quote = FALSE)
#write.csv(freq21_proteins$protein, "data/xls/p1.csv" ,row.names = FALSE, quote = FALSE)
```

### Number of drugs that target proteins at each group

The number of drugs that target proteins at each group of proteins are calculated. Especially, we expect to see that the proteins in freq12 are mainly targeted by drugs in cluster 2 (the similar results for freq21).

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE }
freq21_proteins <-  readRDS(file = "data/rds/freq21_proteins.rds")
freq11_proteins <-  readRDS(file = "data/rds/freq11_proteins.rds")
freq12_proteins <-  readRDS(file = "data/rds/freq12_proteins.rds")
freq22_proteins <-  readRDS(file = "data/rds/freq22_proteins.rds")
df_dtc_fimm1 <- readRDS(file = "data/rds/df_dtc_fimm1.rds")
df_dtc_fimm2 <- readRDS(file = "data/rds/df_dtc_fimm2.rds")
#--------------------------freq12-----------------
freq12_proteins <- left_join(freq12_proteins,df_dtc_fimm1,by = "protein");colnames(freq12_proteins)[colnames(freq12_proteins)=="drug"] <- "drug1"
freq12_proteins <- left_join(freq12_proteins,df_dtc_fimm2,by = "protein");colnames(freq12_proteins)[colnames(freq12_proteins)=="drug"] <- "drug2"
num_drug12_1 <- length(unique(freq12_proteins$drug1));num_drug12_2 <- length(unique(freq12_proteins$drug2))
#---------------------------freq21-------------
freq21_proteins <- left_join(freq21_proteins,df_dtc_fimm1,by = "protein");colnames(freq21_proteins)[colnames(freq21_proteins)=="drug"] <- "drug1"
freq21_proteins <- left_join(freq21_proteins,df_dtc_fimm2,by = "protein");colnames(freq21_proteins)[colnames(freq21_proteins)=="drug"] <- "drug2"
num_drug21_1 <- length(unique(freq21_proteins$drug1));num_drug21_2 <- length(unique(freq21_proteins$drug2))
#-----------------------------freq11---------------
freq11_proteins <- left_join(freq11_proteins,df_dtc_fimm1,by = "protein");colnames(freq11_proteins)[colnames(freq11_proteins)=="drug"] <- "drug1"
freq11_proteins <- left_join(freq11_proteins,df_dtc_fimm2,by = "protein");colnames(freq11_proteins)[colnames(freq11_proteins)=="drug"] <- "drug2"
num_drug11_1 <- length(unique(freq11_proteins$drug1));num_drug11_2 <- length(unique(freq11_proteins$drug2))
#-----------------------------freq22---------------
freq22_proteins <- left_join(freq22_proteins,df_dtc_fimm1,by = "protein");colnames(freq22_proteins)[colnames(freq22_proteins)=="drug"] <- "drug1"
freq22_proteins <- left_join(freq22_proteins,df_dtc_fimm2,by = "protein");colnames(freq22_proteins)[colnames(freq22_proteins)=="drug"] <- "drug2"
num_drug22_1 <- length(unique(freq22_proteins$drug1));num_drug22_2 <- length(unique(freq22_proteins$drug2))

#write.csv(freq11_proteins$protein, "/Users/mehdimir/Documents/Projects/Git/AML_FIMM/Data/xls/p12.csv" ,row.names = FALSE, quote = FALSE) #for GO annotations using https://biit.cs.ut.ee/gprofiler/gost website

```

The proteins in freq12(freq21) analysed using pathway enrichment to show that they are significantly from different pathways.
The enrichment analysis was done using shinyGO and the results saved at folder "xls" under the names BP1.csv, BP2.csv, kegg1.csv,
kegg2.csv. The following four chunks make plots for enrichment analysis results.
## BP1
```{r}
bp1 <- read.csv("data/xls/BP1.csv")
bp1 <- bp1 %>% filter(Enrichment.FDR<0.05 & nGenes > 2)
url_split <- data.frame(do.call("rbind",strsplit(as.character(bp1$URL),"/",fixed = TRUE))) 
bp1 <- bp1 %>% mutate(logfdr = -log(bp1$Enrichment.FDR), go = url_split$X6) 
fdr_go_bp1 <- dplyr::select(bp1,go,Enrichment.FDR)
#write.csv(fdr_go_bp1,"data/xls/GO/fdrgo_bp1.csv", row.names = FALSE)
bp1 <- bp1[c(4,1,21,53,23,13,36,26,121),]
bp1 <- bp1[order(bp1$Fold.Enrichment,decreasing = T), ]  # sort
bp1[2,5] <- "Positive regulation of\n protein import into nucleus "
bp1[5,5] <- "Olefinic compound\n metabolic process" 
#----------------------------visualization---------
bp1f <- ggplot(bp1, aes(x=reorder(Pathway, Fold.Enrichment), y=Fold.Enrichment, color=logfdr)) + 
  geom_point( aes( size = nGenes))  + 
  scale_size(range = c(2, 3.5)) +
  geom_segment(aes(y = 0, 
                   x = reorder(Pathway, Fold.Enrichment), 
                   yend = Fold.Enrichment, 
                   xend = reorder(Pathway, Fold.Enrichment)),size = 1.5) +
    theme_bw() +
    theme(aspect.ratio = 0.82) +
  theme(legend.position = c(0.8, 0.5),legend.title = element_text(size = 8), 
        legend.text = element_text(size = 6),legend.direction = "vertical", 
        legend.box = "horizontal")+
  theme(axis.text.x = element_text(size = 12,colour = "black" ,family="Times New Roman")) +
  theme(axis.text.y = element_text(size = 12,colour = "black" ,family="Times New Roman"))+
  labs( subtitle="C) Biological Process: Protein targets G1", size =12) + 
  guides(size=guide_legend("N. of Genes")) + 
#  scale_color_gradient(low="blue", high="red" , name = "-log10(FDR)") +
  scale_color_viridis_c(name = "-log10(FDR)")+  
  ylab("Fold Enrichment")+
  xlab("") +
  coord_flip()
```

## BP2
```{r}
bp2 <- read.csv("data/xls/BP2.csv")
bp2 <- bp2 %>% filter(Enrichment.FDR<0.05 & nGenes > 2)
url_split <- data.frame(do.call("rbind",strsplit(as.character(bp2$URL),"/",fixed = TRUE))) 
bp2 <- bp2 %>% mutate(logfdr = -log(bp2$Enrichment.FDR), go = url_split$X6) 
fdr_go_bp2 <- dplyr::select(bp2,go,Enrichment.FDR)
#write.csv(fdr_go_bp2,"data/xls/GO/fdrgo_bp2.csv", row.names = FALSE)
bp2 <- bp2[c(1,2,31,42,3,56,11,27,31),]
bp2 <- bp2[order(bp2$Fold.Enrichment,decreasing = T), ]  # sort
bp2[3,5] = "Regulation of cyclin-dependent\n protein kinase activity"
bp2[5,5] = "Protein-containing\n complex assembly"
bp2[8,5] = "Response to oxygen\n containing compound "
#----------------------------visualization---------
bp2f <- ggplot(bp2, aes(x=reorder(Pathway, Fold.Enrichment), y=Fold.Enrichment, color=logfdr)) + 
  geom_point( aes( size = nGenes))  + 
  scale_size(range = c(2, 3.5)) +
  geom_segment(aes(y = 0, 
                   x = reorder(Pathway, Fold.Enrichment), 
                   yend = Fold.Enrichment, 
                   xend = reorder(Pathway, Fold.Enrichment),edge_size = nGenes),size = 1.5) +
  theme_bw() +
   theme(aspect.ratio = 0.82) +
  theme(legend.position = c(0.8, 0.5),legend.title = element_text(size = 8), 
        legend.text = element_text(size = 6),legend.direction = "vertical", 
        legend.box = "horizontal")+
    theme(axis.text.x = element_text(size = 12,colour = "black" ,family="Times New Roman")) +
  theme(axis.text.y = element_text(size = 12,colour = "black" ,family="Times New Roman"))+
  labs( subtitle="D) Biological Process: Protein targets G2", size =12) + 
  guides(size=guide_legend("N. of Genes")) + 
#  scale_color_gradient(low="blue", high="red" , name = "-log10(FDR)") +
  scale_color_viridis_c(name = "-log10(FDR)")+  
  ylab("Fold Enrichment")+
  xlab("") +
  coord_flip()

```

## KEGG1
```{r}
kegg1 <- read.csv("data/xls/kegg1.csv")
kegg1 <- kegg1 %>% filter(Enrichment.FDR<0.05 & nGenes > 4)
#kg1 <- kegg1[c(2,4,6,7,8,12,16),]
kg1 <- kegg1
kg1 <- kg1 %>% mutate(logfdr = -log(kg1$Enrichment.FDR)) 
kg1 <- kg1[order(kg1$Fold.Enrichment,decreasing = T), ]  # sort
#----------------------------visualization---------
kg1f <- ggplot(kg1, aes(x=reorder(Pathway, Fold.Enrichment), y=Fold.Enrichment, color=logfdr)) + 
  geom_point( aes( size = nGenes))  + 
  scale_size(range = c(2, 3.5)) +
  geom_segment(aes(y = 0, 
                   x = reorder(Pathway, Fold.Enrichment), 
                   yend = Fold.Enrichment, 
                   xend = reorder(Pathway, Fold.Enrichment),edge_size = nGenes),size = 1.5) +
  theme_bw() +
  theme(aspect.ratio = 0.82) +
   theme(legend.position = c(0.8, 0.5),legend.title = element_text(size = 8), 
        legend.text = element_text(size = 6),legend.direction = "vertical", 
        legend.box = "horizontal")+
  theme(axis.text.x = element_text(size = 12,colour = "black" ,family="Times New Roman")) +
  theme(axis.text.y = element_text(size = 12,colour = "black" ,family="Times New Roman"))+
  labs( subtitle="E) KEGG: Protein targets G1", size = 12) + 
  guides(size=guide_legend("N. of Genes")) + 
#  scale_color_gradient(low="blue", high="red" , name = "-log10(FDR)") +
  scale_color_viridis_c(name = "-log10(FDR)")+ 
  guides(colour = guide_colourbar(order = 1))+
  ylab("Fold Enrichment")+
  xlab("") +
  coord_flip()
```

## KEGG2
```{r}
kegg2 <- read.csv("data/xls/kegg2.csv")
kegg2 <- kegg2 %>% filter(Enrichment.FDR<0.05 & nGenes > 3)
kg2 <- kegg2 %>% mutate(logfdr = -log(kegg2$Enrichment.FDR)) 
kg2 <- kg2[order(kg2$Fold.Enrichment,decreasing = T), ]  # sort
kg2[4,5] = "Adrenergic signaling\n in cardiomyocytes"
kg2[5,5] = "T-cell leukemia \n virus 1 infection"
#----------------------------visualization---------
kg2f <- ggplot(kg2, aes(x=reorder(Pathway, Fold.Enrichment), y=Fold.Enrichment, color=logfdr)) + 
  geom_point( aes( size = nGenes))  + 
  scale_size(range = c(2, 3.5)) +
  geom_segment(aes(y = 0, 
                   x = reorder(Pathway, Fold.Enrichment), 
                   yend = Fold.Enrichment, 
                   xend = reorder(Pathway, Fold.Enrichment),edge_size = nGenes),size = 1.5) +
  theme_bw() +
     theme(aspect.ratio = 0.82) +
  theme(legend.position = c(0.8, 0.5),legend.title = element_text(size = 8), 
        legend.text = element_text(size = 6),legend.direction = "vertical", 
        legend.box = "horizontal")+
  theme(axis.text.y = element_text(size = 12 ,colour = "black" ,family="Times New Roman")) +
  theme(axis.text.x = element_text(size = 12 ,colour = "black" , family="Times New Roman"))+
  labs(subtitle="F) KEGG: Protein targets G2",size = 12) + 
  guides(size=guide_legend("N. of Genes")) + 
#  scale_color_gradient(low="blue", high="red" , name = "-log10(FDR)") +
  scale_color_viridis_c(name = "-log10(FDR)")+  
  ylab("Fold Enrichment")+
  xlab("") +
  coord_flip()

```

 The following chunk make a grid of four BP1, BP2, kegg1, kegg2 plots
```{r}
library("cowplot")
library("grid")

merg <- plot_grid(bp1f, bp2f,kg1f,kg2f,
          ncol = 2,
          labels = c("A","B","C","D"))
figure <- ggarrange(bp1f, bp2f,kg1f,kg2f,
                    labels = c("", "","",""),
                    ncol = 2, nrow = 2)
figure

#save 16x 16
```

##  DICE similarity if drugs at each cluster

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(rcdk)
dtc <- readRDS(file = "data/rds/dtc.rds")
drug_inchi <- readRDS(file = "data/rds/drug_inchi.rds")
drug_inchi_dtc <- left_join(drug_inchi,dtc,by = "InChI.Key")
drug_inchi_dtc <- drug_inchi_dtc[,c(1,2)]
colnames(drug_inchi_dtc) <- c("drug","InChI.Key")
drug_inchi_dtc <- drug_inchi_dtc[!duplicated(drug_inchi_dtc), ]##removing duplicate rows.
#-----------------reading smile file----------------
smiles <- read.csv("data/xls/smiles.csv", sep=";")
colnames(smiles) <- c("InChI.Key","Isomeric_smile","Canonical_smile")
smile_can <- smiles[,c(1,3)]
drug_inchi_smile <- left_join(drug_inchi_dtc,smile_can,by = "InChI.Key") %>% dplyr::filter(Canonical_smile !="")
#------------------------------------------------------
g <- readRDS(file = "data/rds/g.rds")
drug1 <- as.character(V(g)$name[V(g)$memb == 1])
drug2 <- as.character(V(g)$name[V(g)$memb == 2])
smile_cl1 <- drug_inchi_smile %>% filter(drug  %in%  drug1)
smile_cl2 <- drug_inchi_smile %>% filter(drug  %in%  drug2)
mols1 <- parse.smiles(smile_cl1[ , 3, drop = TRUE])
fps1 <- lapply(mols1, get.fingerprint, type='graph', size = 1024)
fp.sim1 <- fingerprint::fp.sim.matrix(fps1, method = 'dice')
mols2 <- parse.smiles(smile_cl2[ , 3, drop = TRUE])
fps2 <- lapply(mols2, get.fingerprint, type='graph', size = 1024)
fp.sim2 <- fingerprint::fp.sim.matrix(fps2, method = 'dice')
bet_cl <- matrix(NA, nrow = length(fps1), ncol = length(fps2))
for (i in 1:length(fps1)){
  fp1 <- fps1[[i]]
  for (j in 1:length(fps2)){
    fp2 <- fps2[[j]]
    bet_cl[i,j] <- fingerprint::fp.sim.matrix( list(fp1,fp2) )[1,2]
  }
}
rand12 <- as.vector(bet_cl)

diag(fp.sim1) <- 0
clust1_sim <- t(fp.sim1)[lower.tri(t(fp.sim1))]
# 2nd cluster  
diag(fp.sim2) <- 0
clust2_sim <- t(fp.sim2)[lower.tri(t(fp.sim2))]

#Test significant difference between cluster 1 and 2.
set.seed(5)
ind_rnd <- sample (c(1:18189), size=10000, replace=F)
dat <- data.frame(Cluster =
                  factor(c(rep("Cluster 1",length(clust1_sim)),
                           rep("Cluster 2",length(clust2_sim)),
                           rep("Intercluster",10000))), 
                  sim = c(clust1_sim,clust2_sim,rand12[ind_rnd]))
head(dat)
# Density plots with semi-transparent fill

library(plyr)
library(ggpubr)
cdat <- ddply(dat, "Cluster", summarise, sim.mean=mean(sim))
ggplot(dat, aes(x=sim, colour=Cluster)) +
    geom_density() +
    geom_vline(data=cdat, aes(xintercept=sim.mean,  colour=Cluster),
               linetype="dashed", size=1)

ggplot(dat, aes(x=Cluster, y=sim, fill=Cluster,color =  Cluster)) + 
  geom_boxplot(alpha = 0.5, col = "gray40") +   
  theme_pubr(border = TRUE)+
#  stat_summary(fun.x="mean")+
  theme(legend.position="top",legend.text = element_text(size=10, face="bold"),
        legend.title=element_blank()) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  xlab("")+
  ylab("Dice similarity of SMILEs") +
  scale_fill_manual(values = c("slateblue3", "sienna1", "#353436")) +
  coord_flip()
#save 8x 4 pdf


x <- c(clust1_sim,rand12)
g <- factor(c(rep("1",length(clust1_sim)),rep("12",length(rand12))))
kruskal.test(x, g)
x <- c(clust2_sim,rand12)
g <- factor(c(rep("2",length(clust2_sim)),rep("12",length(rand12))))
kruskal.test(x, g)
x <- c(clust1_sim,clust2_sim)
g <- factor(c(rep("1",length(clust1_sim)),rep("2",length(clust2_sim))))


```

# Drug selection

## Efficacy&Toxicity calculatin

The average of efficacy and toxicity of each drug on 81 samples are calculated.

```{r}
sample_drug_response_healthy <- readRDS("data/rds/sample_drug_response_healthy.rds")
max_rec_mean <- readRDS("data/rds/max_rec_mean.rds")
max_rec_minmax <- readRDS("data/rds/max_rec_minmax.rds")
g <- readRDS(file = "data/rds/g.rds")
toxicity <- sample_drug_response_healthy %>% group_by(drug) %>% summarise(tox = mean(mean)) 
w_toxicity <- sample_drug_response_healthy %>% group_by(drug) %>% summarise(w_tox = mean(minmax)) 
tox_wtox <- left_join(toxicity,w_toxicity,by="drug") 
effi <- as.data.frame(apply(max_rec_mean, 2,mean));w_effi <- as.data.frame(apply(max_rec_minmax, 2,mean))
drug = V(g)$name
drug_member_mean <- data_frame(drug = V(g)$name, 
                               member = as.factor(V(g)$memb), 
                               efficacy = effi[drug,],
                               w_efficacy = w_effi[drug,])
drug_mem_eff_tox <- left_join(drug_member_mean,tox_wtox,by="drug")
#drug_mem_eff_tox <- saveRDS(drug_mem_eff_tox,"data/rds/drug_mem_eff_tox.rds")
```

## drug selection based on P-value < 0.05

```{r}
drug_mem_eff_tox <- readRDS(file="data/rds/drug_mem_eff_tox.rds")
pval <- rep(100, 296)
mue <- mean(drug_mem_eff_tox$efficacy);mut <- mean(drug_mem_eff_tox$tox)
for (i in 1:296){
  efi <- drug_mem_eff_tox$efficacy[i];toxi <- drug_mem_eff_tox$tox[i]
  if (efi>mue & toxi < mut ){  
  Fi <- drug_mem_eff_tox %>% filter(drug_mem_eff_tox$efficacy >= efi & drug_mem_eff_tox$tox <=toxi)
  pval [i] <- dim(Fi)[1]/296
  }
}
drug_mem_eff_tox_pval<- cbind(drug_mem_eff_tox,pval)
drug_selected <- drug_mem_eff_tox %>% dplyr::filter(pval < 0.05)
#saveRDS(drug_mem_eff_tox_pval,file = "data/rds/drug_mem_eff_tox_pval.rds")
#saveRDS(drug_selected, file = "data/rds/drug_selected.rds")
```

## Visualization

```{r}
library(cowplot)
drug_mem_eff_tox <- readRDS(file = "data/rds/drug_mem_eff_tox_pval.rds")

main.plot <- ggplot(drug_mem_eff_tox, aes(x = efficacy, y = tox))+
  geom_density_2d_filled(alpha = 0.5) +
  geom_density_2d(size = 0.25, colour = "black")+
  geom_point(aes(x = efficacy, y = tox, col = -log(pval)))  +
  theme_bw() +
  theme(legend.position = "none",axis.title = element_blank()) 
#  xlab("Inhibition rate based on AMl samples") + ylab("Inhibition based on Healthy sample")
inset.plot <- ggplot(drug_mem_eff_tox %>% filter(efficacy <=15 & tox <=15), aes(x = efficacy, y = tox))+
  geom_density_2d_filled(alpha = 0.5) +
  geom_density_2d(size = 0.25, colour = "black")+
  geom_point(aes(x = efficacy, y = tox, col = -log(pval)))+
  geom_label_repel(data=filter(drug_mem_eff_tox, pval<0.05), size =4, fontface = "bold",alpha = 0.7, aes(label=drug), max.overlaps = 30)+
  theme_bw() +
  theme(legend.position = "none", axis.title = element_text(size = 12)  ) +
  xlab("Percentage of inhibition based on AMl samples (Efficacy)") + ylab("Percentage of inhibition based on Healthy sample (Toxicity)")
plot.with.inset <-
  ggdraw() +
  draw_plot(inset.plot) +
  draw_plot(main.plot, x = 0.06, y = .55, width = .45, height = .45)
#save 8x6 pdf
ggplot(drug_mem_eff_tox, aes(x = efficacy, y = tox))+
  geom_density_2d_filled(alpha = 0.5) +
  geom_density_2d(size = 0.25, colour = "black")+
  geom_point(aes(x = efficacy, y = tox, col = -log(pval)))+
  geom_label_repel(data=filter(drug_mem_eff_tox, pval<0.05), size =2, fontface = "bold",alpha = 0.7, aes(label=drug), max.overlaps = 30)+
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Efficacy") + ylab("Toxicity")
```

# CTG data analysis

The synergy of selected drug combinations tested in CTG readout is calculated in this section.
We have two folders AML and healthy include the response of drug combinations on diagonal matrix.
We run decrease model to construct the whole matrix and then calculate the synergy scoring.
We do not need to run this chunk. If you want to run this chunk you need to uncomment line 938 and 956 for healthy samples and also lines 939 and 955 for AML samples.

## Decrease model on CTG results
```{r}
library(synergyfinder)
library(readxl)
library(dplyr)
drug_selected <- readRDS("data/rds/drug_selected.rds")
dr1 <- drug_selected %>% filter(member ==1) %>% dplyr::select(drug)
dr2 <- drug_selected %>% filter(member ==2) %>% dplyr::select(drug)
files <-list.files("data/xls/CTG/healthy/")
#files <-list.files("data/xls/CTG/AML/")
k <- 1
n <- length(files)*28
htb_res<- data.frame(sample = character(n),drug1 = character(n),
                     drug2=character(n),group=character(n),
                     cluster=character(n),mu1=double(n),
                     high_dose1=double(n),mu2=double(n),high_dose2=double(n),
                     mu=double(n),mu_diag = double(n),high_dose = double(n),
                     md=double(n) ,md_diag=double(n),
                     IR=double(n) ,IR_hd=double(n) , IR_diag=double(n),
                     zip=double(n),zip_diag=double(n),zip_high_dose=double(n),
                     hsa=double(n),hsa_diag=double(n),hsa_high_dose=double(n),
                     bliss=double(n),bliss_diag=double(n),bliss_high_dose=double(n),
                     loewe=double(n),loewe_diag=double(n),loewe_high_dose=double(n))
for (f in files) {
 # f <- files[5]
 # g <- paste0("data/xls/CTG/AML/", f)
  g <- paste0("data/xls/CTG/healthy/", f)
  decrease <- read_excel(g)
  pos_result <- decrease %>% 
    filter((Drug1 %in% dr1$drug & Drug2 %in% dr2$drug)
           |(Drug1 %in% dr2$drug & Drug2 %in% dr1$drug))
  neg_result <- decrease %>% 
    filter(!(Drug1 %in% dr1$drug & Drug2 %in% dr2$drug)
           |(Drug1 %in% dr2$drug & Drug2 %in% dr1$drug))
  inds_pos <- unique(pos_result$PairIndex)
#---------------------------------------------------------
  for (i in inds_pos){
#    i= 4
    datai <- filter(pos_result,PairIndex == i)
    datai$Response[datai$Response <0 ] <- 0
    datai$Response[datai$Response > 100 ] <- 100
    data <- ReshapeData(datai,data_type = "inhibition")
    data$drug_pairs$conc_unit2 = 'nM'
    data$drug_pairs$conc_unit1 = 'nM'
    datax <- CalculateSynergy(data)
    htb_res$sample[k] <-f
    htb_res$group[k] <-"Positive"
    htb_res$drug1[k] <-datax$drug_pairs$drug1
    htb_res$drug2[k] <-datax$drug_pairs$drug2
    if (htb_res$drug1[k] %in% dr1$drug & htb_res$drug2[k] %in% dr1$drug){
      htb_res$cluster[k] <- "11"
      }else if (htb_res$drug1[k] %in% dr2$drug & htb_res$drug2[k] %in% dr2$drug){
      htb_res$cluster[k] = "22"
      }else{
      htb_res$cluster[k] =  "12"
      }
    #----------inhibition---------------------
    #-----------single------------------------
    data1i <- datai %>% filter((Conc2 == 0))%>% filter(!(Conc1 == 0))
    htb_res$mu1[k] <- mean(data1i$Response)
    htb_res$high_dose1[k] <- mean(data1i$Response[5])
    data2i <- datai %>% filter((Conc1 == 0))%>% filter(!(Conc2 == 0))
    htb_res$mu2[k] <- mean(data2i$Response)
    htb_res$high_dose2[k] <- mean(data2i$Response[5])
    #--------------combinaion------------------
    datai2 <- datai %>% filter(!(Conc1 == 0 | Conc2 == 0)) 
    htb_res$mu[k] <- mean(datai2$Response)
    mat_inh <- datai2[,c("Conc1","Conc2","Response")] %>%
      spread(key = Conc1, value = Response) %>% as.matrix()
    rownames(mat_inh) <- mat_inh[,1];mat_inh <- mat_inh[,-1]
    htb_res$mu_diag[k] <- mean(mat_inh[col(mat_inh)==row(mat_inh)])
    htb_res$high_dose[k] <- mat_inh[5,5]
    htb_res$md[k] = median(datai2$Response)
    htb_res$md_diag[k]= median(mat_inh[col(mat_inh)==row(mat_inh)])
    #----------------IR & Idiff-------------------
    htb_res$IR[k] = htb_res$mu[k]/max(htb_res$mu1[k],htb_res$mu2[k])
    htb_res$IR_hd[k] = htb_res$high_dose[k]/max(htb_res$high_dose1[k],htb_res$high_dose2[k])
    htb_res$IR_diag[k]= htb_res$mu_diag[k]/max(htb_res$mu1[k],htb_res$mu2[k])
    #------------zip synergy-----------------------
    syn <- datax$synergy_scores %>% filter(!(conc1 == 0 | conc2 == 0)) 
    htb_res$zip[k] <-  mean(syn$ZIP_synergy)
    mat_zip <- syn[,c("conc1","conc2","ZIP_synergy")] %>%
      spread(key = conc1, value = ZIP_synergy) %>% as.matrix()
    rownames(mat_zip) <- mat_zip[,1];mat_zip <- mat_zip[,-1]
    htb_res$zip_diag[k] <- mean(mat_zip[col(mat_zip)==row(mat_zip)])
    htb_res$zip_high_dose[k] <- mat_zip[5,5]
    #------------hsa synergy-----------------------
    htb_res$hsa[k] <-  mean(syn$HSA_synergy)
    mat_hsa <- syn[,c("conc1","conc2","HSA_synergy")] %>%
      spread(key = conc1, value = HSA_synergy) %>% as.matrix()
    rownames(mat_hsa) <- mat_hsa[,1];mat_hsa <- mat_hsa[,-1]
    htb_res$hsa_diag[k] <- mean(mat_hsa[col(mat_hsa)==row(mat_hsa)])
    htb_res$hsa_high_dose[k] <- mat_hsa[5,5]
   #------------bliss synergy-----------------------
    htb_res$bliss[k] <-  mean(syn$Bliss_synergy)
    mat_bliss <- syn[,c("conc1","conc2","Bliss_synergy")] %>%
      spread(key = conc1, value = Bliss_synergy) %>% as.matrix()
    rownames(mat_bliss) <- mat_bliss[,1];mat_bliss <- mat_bliss[,-1]
    htb_res$bliss_diag[k] <- mean(mat_bliss[col(mat_bliss)==row(mat_bliss)])
    htb_res$bliss_high_dose[k] <- mat_bliss[5,5]
    #------------loewe synergy-----------------------
    htb_res$loewe[k] <-  mean(syn$Loewe_synergy)
    mat_loewe <- syn[,c("conc1","conc2","Loewe_synergy")] %>%
      spread(key = conc1, value = Loewe_synergy) %>% as.matrix()
    rownames(mat_loewe) <- mat_loewe[,1];mat_loewe <- mat_loewe[,-1]
    htb_res$loewe_diag[k] <- mean(mat_loewe[col(mat_loewe)==row(mat_loewe)])
    htb_res$loewe_high_dose[k] <- mat_loewe[5,5]
    k <- k+1
  }
  inds_neg <- unique(neg_result$PairIndex)
  for (i in inds_neg){
    datai <- filter(neg_result,PairIndex == i)
    datai$Response[datai$Response <0 ] <- 0
    datai$Response[datai$Response > 100 ] <- 100
    data <- ReshapeData(datai,data_type = "inhibition")
    data$drug_pairs$conc_unit2 = 'nM'
    data$drug_pairs$conc_unit1 = 'nM'
    datax <- CalculateSynergy(data)
    htb_res$sample[k] <-f
    htb_res$group[k] <-"Negative"
    htb_res$drug1[k] <-datax$drug_pairs$drug1
    htb_res$drug2[k] <-datax$drug_pairs$drug2
    if (htb_res$drug1[k] %in% dr1$drug & htb_res$drug2[k] %in% dr1$drug){
      htb_res$cluster[k] <- "11"}
    else if (htb_res$drug1[k] %in% dr2$drug & htb_res$drug2[k] %in% dr2$drug){
      htb_res$cluster[k] = "22"
      }else{
      htb_res$cluster[k] =  "12"
      }
    #----------inhibition--------------
    data1i <- datai %>% filter((Conc2 == 0))%>% filter(!(Conc1 == 0))
    htb_res$mu1[k] <- mean(data1i$Response)
    htb_res$high_dose1[k] <- mean(data1i$Response[5])
    data2i <- datai %>% filter((Conc1 == 0))%>% filter(!(Conc2 == 0))
    htb_res$mu2[k] <- mean(data2i$Response)
    htb_res$high_dose2[k] <- mean(data2i$Response[5])

    datai2 <- datai %>% filter(!(Conc1 == 0 | Conc2 == 0)) 
    htb_res$mu[k] <- mean(datai2$Response)
    mat_inh <- datai2[,c("Conc1","Conc2","Response")] %>%
      spread(key = Conc1, value = Response) %>% as.matrix()
    rownames(mat_inh) <- mat_inh[,1];mat_inh <- mat_inh[,-1]
    htb_res$mu_diag[k] <- mean(mat_inh[col(mat_inh)==row(mat_inh)])
    htb_res$high_dose[k] <- mat_inh[5,5]
    htb_res$md[k] = median(datai2$Response)
    htb_res$md_diag[k]= median(mat_inh[col(mat_inh)==row(mat_inh)])
    #----------------IR & Idiff-------------------
    htb_res$IR[k] = htb_res$mu[k]/max(htb_res$mu1[k],htb_res$mu2[k])
    htb_res$IR_hd[k] = htb_res$high_dose[k]/max(htb_res$high_dose1[k],htb_res$high_dose2[k])
    htb_res$IR_diag[k]= htb_res$mu_diag[k]/max(htb_res$mu1[k],htb_res$mu2[k])
    #------------zip synergy-----------------------
    syn <- datax$synergy_scores %>% filter(!(conc1 == 0 | conc2 == 0)) 
    htb_res$zip[k] <-  mean(syn$ZIP_synergy)
    mat_zip <- syn[,c("conc1","conc2","ZIP_synergy")] %>%
      spread(key = conc1, value = ZIP_synergy) %>% as.matrix()
    rownames(mat_zip) <- mat_zip[,1];mat_zip <- mat_zip[,-1]
    htb_res$zip_diag[k] <- mean(mat_zip[col(mat_zip)==row(mat_zip)])
    htb_res$zip_high_dose[k] <- mat_zip[5,5]
    #------------hsa synergy-----------------------
    htb_res$hsa[k] <-  mean(syn$HSA_synergy)
    mat_hsa <- syn[,c("conc1","conc2","HSA_synergy")] %>%
      spread(key = conc1, value = HSA_synergy) %>% as.matrix()
    rownames(mat_hsa) <- mat_hsa[,1];mat_hsa <- mat_hsa[,-1]
    htb_res$hsa_diag[k] <- mean(mat_hsa[col(mat_hsa)==row(mat_hsa)])
    htb_res$hsa_high_dose[k] <- mat_hsa[5,5]
   #------------bliss synergy-----------------------
    htb_res$bliss[k] <-  mean(syn$Bliss_synergy)
    mat_bliss <- syn[,c("conc1","conc2","Bliss_synergy")] %>%
      spread(key = conc1, value = Bliss_synergy) %>% as.matrix()
    rownames(mat_bliss) <- mat_bliss[,1];mat_bliss <- mat_bliss[,-1]
    htb_res$bliss_diag[k] <- mean(mat_bliss[col(mat_bliss)==row(mat_bliss)])
    htb_res$bliss_high_dose[k] <- mat_bliss[5,5]
    #------------loewe synergy-----------------------
    htb_res$loewe[k] <-  mean(syn$Loewe_synergy)
    mat_loewe <- syn[,c("conc1","conc2","Loewe_synergy")] %>%
      spread(key = conc1, value = Loewe_synergy) %>% as.matrix()
    rownames(mat_loewe) <- mat_loewe[,1];mat_loewe <- mat_loewe[,-1]
    htb_res$loewe_diag[k] <- mean(mat_loewe[col(mat_loewe)==row(mat_loewe)])
    htb_res$loewe_high_dose[k] <- mat_loewe[5,5]
    k <- k+1
  }
} 
#saveRDS(htb_healthy,file = "data/rds/ctg_healthy.rds")
#saveRDS(htb_aml,file = "data/rds/ctg_aml.rds")
```

## Toxicity vs Efficacy plots
The following 6 chunk, make plots of response and synergy values for AML and healthy samples
### mean response
```{r}
library(rstatix)
ctg_healthy <- readRDS(file  = "data/rds/ctg_healthy.rds")
ctg_aml <- readRDS(file = "data/rds/ctg_aml.rds")
ctg_aml$drug1 = substr(ctg_aml$drug1, 1, 4)
ctg_aml$drug2 = substr(ctg_aml$drug2, 1, 4)
ctg_healthy$drug1 = substr(ctg_healthy$drug1, 1, 4)
ctg_healthy$drug2 = substr(ctg_healthy$drug2, 1, 4)
heal_com <- ctg_healthy %>% dplyr::select(sample,drug1,drug2,mu_diag) %>% 
  group_by(drug1,drug2) %>% summarise(mu_diag_heal=mean(mu_diag))
aml_com <- ctg_aml %>% dplyr::select(sample,drug1,drug2,cluster,group,mu_diag) %>%
  group_by(drug1,drug2,cluster,group) %>% summarise(mu_diag_aml=mean(mu_diag))
ctg_aml_hel <- left_join(aml_com,heal_com, by = c("drug1","drug2"))%>% 
  mutate(ratio = mu_diag_aml/mu_diag_heal)
#--------------------------P-value--------------------------------
data <- ctg_aml_hel %>% filter(mu_diag_aml > quantile(ctg_aml_hel$mu_diag_aml,probs =0.25) & mu_diag_heal<quantile(ctg_aml_hel$mu_diag_heal,probs =0.75))
prop_test(x = 9, n = 14, p = 0.3, detailed = TRUE,alternative = "greater")
#------------------------------------------------------------------
g_diag <- ggplot(ctg_aml_hel, aes(x = mu_diag_aml, y = mu_diag_heal)) +
  geom_point(aes(colour = cluster), size = 5, alpha = 0.6) +
  geom_label_repel(data=filter(ctg_aml_hel, mu_diag_aml> median(ctg_aml_hel$mu_diag_aml) & 
                                 mu_diag_heal < median(ctg_aml_hel$mu_diag_heal)), 
                   size =4.5, fontface = "bold",alpha = 0.7, 
                   aes(label=paste0(drug1 ,",", drug2)), 
                   max.overlaps = Inf,nudge_x = -1, 
                   box.padding = 0.5,nudge_y = -1,
                   segment.curvature = -0.1,segment.angle = 20) +
 geom_hline(yintercept = median(ctg_aml_hel$mu_diag_heal), linetype = "dashed", color = "blue") +
 geom_vline(xintercept = median(ctg_aml_hel$mu_diag_aml), linetype = "dashed", color = "red")+
 annotate("text", x=46, y=18, label= "P-value < 0.006", size = 4.5, fontface = "bold",color = "black") +
 theme(panel.background = element_rect(fill = "white", colour = "grey50")) +
 theme(legend.position = c(0.15, 0.82),legend.title = element_text(size = 8), 
               legend.text = element_text(size = 8)) +
 theme(axis.text.y = element_text(face = "bold",size = 14)) +
 theme(axis.text.x = element_text(face = "bold",size = 14)) +
 xlab("Efficacy (AML Samples)") + ylab("Toxicity (Healthy Samples)") +
 theme(axis.title.x = element_text(face = "bold",size = 14)) +
 theme(axis.title.y = element_text(face = "bold",size = 14)) +
scale_color_manual(name = "",labels=c("Cluster 1","Intercluster","Cluster 2"),values = c("#ffcf20FF","#10a53dFF","#541352FF")) +
  theme(legend.text = element_text(size = 8, face = "bold",colour = "black"))

```

### zip
```{r}
ctg_healthy <- readRDS(file  = "data/rds/ctg_healthy.rds")
ctg_aml <- readRDS(file = "data/rds/ctg_aml.rds")
ctg_aml$drug1 = substr(ctg_aml$drug1, 1, 4)
ctg_aml$drug2 = substr(ctg_aml$drug2, 1, 4)
ctg_healthy$drug1 = substr(ctg_healthy$drug1, 1, 4)
ctg_healthy$drug2 = substr(ctg_healthy$drug2, 1, 4)
heal_com_hd <- ctg_healthy %>% dplyr::select(sample,drug1,drug2,zip_high_dose) %>% group_by(drug1,drug2) %>% summarise(zip_heal_hd=mean(zip_high_dose))
aml_com_hd <- ctg_aml %>% dplyr::select(sample,drug1,drug2,cluster,group,zip_high_dose) %>%
  filter(zip_high_dose>0) %>% group_by(drug1,drug2,cluster,group) %>% summarise(zip_aml_hd=mean(zip_high_dose))
ctg_aml_hel <- left_join(aml_com_hd,heal_com_hd, by = c("drug1","drug2")) %>% 
  mutate(ratio = zip_aml_hd/zip_heal_hd)
#--------------------------P-value--------------------------------
data <- ctg_aml_hel %>% filter(zip_aml_hd > quantile(ctg_aml_hel$zip_aml_hd,probs =0.25) & zip_heal_hd<quantile(ctg_aml_hel$zip_heal_hd,probs =0.75))
prop_test(x = 8, n = 13, p = 0.3, detailed = TRUE,alternative = "greater")
#------------------------------------------------------------------
g_zip_hd <- ggplot(ctg_aml_hel, aes(x = zip_aml_hd, y = zip_heal_hd)) +
  geom_point(aes(colour = cluster), size = 5, alpha = 0.6) +
  geom_label_repel(data=filter(ctg_aml_hel, zip_aml_hd> median(ctg_aml_hel$zip_aml_hd) & 
                                 zip_heal_hd < median(ctg_aml_hel$zip_heal_hd)), 
                   size =4.5, fontface = "bold",alpha = 0.7, 
                   aes(label=paste0(drug1 ,",", drug2)), 
                   nudge_x = .2, 
                   box.padding = 0.5,nudge_y = 1,
                   segment.curvature = -0.1,segment.angle = 20) +
 geom_hline(yintercept = median(ctg_aml_hel$zip_heal_hd), linetype = "dashed", color = "blue") +
 geom_vline(xintercept = median(ctg_aml_hel$zip_aml_hd), linetype = "dashed", color = "red")+
 annotate("text", x=11, y=-16, label= "P-value < 0.01", size = 4.5, fontface = "bold") +
 theme(panel.background = element_rect(fill = "white", colour = "grey50"))+
 theme(legend.position = c(0.15, 0.82),legend.title = element_text(size = 8), 
               legend.text = element_text(size = 8))+
 theme(axis.text.y = element_text(face = "bold",size = 14)) +
 theme(axis.text.x = element_text(face = "bold",size = 14)) +
 xlab("ZIP (AML Samples)") + ylab("ZIP (Healthy Samples)") +
theme(axis.title.x = element_text(face = "bold",size = 14)) +
 theme(axis.title.y = element_text(face = "bold",size = 14)) +
 scale_color_manual(name = "",labels=c("Cluster 1","Intercluster","Cluster 2"),values = c("#ffcf20FF","#10a53dFF","#541352FF"))  +
 theme(legend.text = element_text(size = 8, face = "bold"))

#ggsave(paste0("data/wetlab/fig/tix_eff/ziphd.png"),width = 10, height = 15)
```

### hsa
```{r}
ctg_healthy <- readRDS(file  = "data/rds/ctg_healthy.rds")
ctg_aml <- readRDS(file = "data/rds/ctg_aml.rds")
ctg_aml$drug1 = substr(ctg_aml$drug1, 1, 4)
ctg_aml$drug2 = substr(ctg_aml$drug2, 1, 4)
ctg_healthy$drug1 = substr(ctg_healthy$drug1, 1, 4)
ctg_healthy$drug2 = substr(ctg_healthy$drug2, 1, 4)
heal_com_hd <- ctg_healthy %>% dplyr::select(sample,drug1,drug2,hsa_high_dose) %>% group_by(drug1,drug2) %>% summarise(hsa_heal_hd=mean(hsa_high_dose))
aml_com_hd <- ctg_aml %>% dplyr::select(sample,drug1,drug2,cluster,group,hsa_high_dose) %>%
  filter(hsa_high_dose>0) %>% group_by(drug1,drug2,cluster,group) %>% summarise(hsa_aml_hd=mean(hsa_high_dose))
ctg_aml_hel <- left_join(aml_com_hd,heal_com_hd, by = c("drug1","drug2")) %>% 
 mutate(ratio = hsa_aml_hd/hsa_heal_hd)
#-----------------------P-value--------------------------
data <- ctg_aml_hel %>% filter(hsa_aml_hd > quantile(ctg_aml_hel$hsa_aml_hd,probs =0.25) & hsa_heal_hd<quantile(ctg_aml_hel$hsa_heal_hd,probs =0.75))
prop_test(x = 8, n = 14, p = 0.3, detailed = TRUE,alternative = "greater")
#---------------------------------------------------------
g_hsa_hd <- ggplot(ctg_aml_hel, aes(x = hsa_aml_hd, y = hsa_heal_hd)) +
  geom_point(aes(colour = cluster), size = 5, alpha = 0.6) +
  geom_label_repel(data=filter(ctg_aml_hel, hsa_aml_hd> mean(ctg_aml_hel$hsa_aml_hd) & 
                                 hsa_heal_hd < median(ctg_aml_hel$hsa_heal_hd)), 
                   size =4.5, fontface = "bold",alpha = 0.7, 
                   aes(label=paste0(drug1 ,",", drug2)), 
                   max.overlaps = Inf,nudge_x = .2, 
                   box.padding = 0.5,nudge_y = 1,
                   segment.curvature = -0.1,segment.angle = 20) +
 geom_hline(yintercept = median(ctg_aml_hel$hsa_heal_hd), linetype = "dashed", color = "blue") +
 geom_vline(xintercept = mean(ctg_aml_hel$hsa_aml_hd), linetype = "dashed", color = "red")+    
 annotate("text", x=12, y=-2, label= "P-value < 0.02",  size = 4.5, fontface = "bold") +
 theme(panel.background = element_rect(fill = "white", colour = "grey50"))+
 theme(legend.position = c(0.15, 0.82),legend.title = element_text(size = 8), 
               legend.text = element_text(size = 8))+
 theme(axis.text.y = element_text(face = "bold",size = 14)) +
 theme(axis.text.x = element_text(face = "bold",size = 14)) +
theme(axis.title.x = element_text(face = "bold",size = 14)) +
 theme(axis.title.y = element_text(face = "bold",size = 14)) +
scale_color_manual(name = "",labels=c("Cluster 1","Intercluster","Cluster 2"),values = c("#ffcf20FF","#10a53dFF","#541352FF")) +
  theme(legend.text = element_text(size = 8, face = "bold")) +
 xlab("hsa (AML Samples)") + ylab("hsa (Healthy Samples)") 

#ggsave(paste0("data/wetlab/fig/tix_eff/hsahd.png"),width = 10, height = 15)
```

### bliss
```{r} 
ctg_healthy <- readRDS(file  = "data/rds/ctg_healthy.rds")
ctg_aml <- readRDS(file = "data/rds/ctg_aml.rds")
ctg_aml$drug1 = substr(ctg_aml$drug1, 1, 4)
ctg_aml$drug2 = substr(ctg_aml$drug2, 1, 4)
ctg_healthy$drug1 = substr(ctg_healthy$drug1, 1, 4)
ctg_healthy$drug2 = substr(ctg_healthy$drug2, 1, 4)
heal_com_hd <- ctg_healthy %>% dplyr::select(sample,drug1,drug2,bliss_high_dose) %>% group_by(drug1,drug2) %>% summarise(bliss_heal_hd=mean(bliss_high_dose))
aml_com_hd <- ctg_aml %>% dplyr::select(sample,drug1,drug2,cluster,group,bliss_high_dose) %>%
  filter(bliss_high_dose>0) %>% group_by(drug1,drug2,cluster,group) %>% summarise(bliss_aml_hd=mean(bliss_high_dose))
ctg_aml_hel <- left_join(aml_com_hd,heal_com_hd, by = c("drug1","drug2")) %>% 
 mutate(ratio = bliss_aml_hd/bliss_heal_hd)
#------------------P-value--------------------------------
data <- ctg_aml_hel %>% filter(bliss_aml_hd > quantile(ctg_aml_hel$bliss_aml_hd,probs =0.25) & bliss_heal_hd<quantile(ctg_aml_hel$bliss_heal_hd,probs =0.75))
prop_test(x = 9, n = 14, p = 0.3, detailed = TRUE,alternative = "greater")
#-----------------------------------------------------------------------
g_bliss_hd <- ggplot(ctg_aml_hel, aes(x = bliss_aml_hd, y = bliss_heal_hd)) +
  geom_point(aes(colour = cluster), size = 5, alpha = 0.6) +
  geom_label_repel(data=filter(ctg_aml_hel, bliss_aml_hd> median(ctg_aml_hel$bliss_aml_hd) & 
                                 bliss_heal_hd < median(ctg_aml_hel$bliss_heal_hd)), 
                   size =4.5, fontface = "bold",alpha = 0.7, 
                   aes(label=paste0(drug1 ,",", drug2)), 
                   box.padding = 0.5,
                    ylim=c(-12,10),
                   xlim=c(0,16),
                   segment.curvature = -0.1,segment.angle = 20) +
 geom_hline(yintercept = median(ctg_aml_hel$bliss_heal_hd), linetype = "dashed", color = "blue") +
 geom_vline(xintercept = median(ctg_aml_hel$bliss_aml_hd), linetype = "dashed", color = "red")+
 annotate("text", x=11, y=-12, label= "P-value < 0.006", size = 4.5, fontface = "bold") +
 theme(panel.background = element_rect(fill = "white", colour = "grey50"))+
 theme(legend.position = c(0.15, 0.82),legend.title = element_text(size = 8), 
               legend.text = element_text(size = 8,face = "bold"))+
 theme(axis.text.y = element_text(face = "bold",size = 14)) +
 theme(axis.text.x = element_text(face = "bold",size = 14)) +
theme(axis.title.x = element_text(face = "bold",size = 14)) +
 theme(axis.title.y = element_text(face = "bold",size = 14)) +
 scale_color_manual(name = "",labels=c("Cluster 1","Intercluster","Cluster 2"),values = c("#ffcf20FF","#10a53dFF","#541352FF")) +
 xlab("bliss (AML Samples)") + ylab("bliss (Healthy Samples)") 

#ggsave(paste0("data/wetlab/fig/tix_eff/blisshd.png"),width = 10, height = 15)
```

### loewe
```{r}
ctg_healthy <- readRDS(file  = "data/rds/ctg_healthy.rds")
ctg_aml <- readRDS(file = "data/rds/ctg_aml.rds")
ctg_aml$drug1 = substr(ctg_aml$drug1, 1, 4)
ctg_aml$drug2 = substr(ctg_aml$drug2, 1, 4)
ctg_healthy$drug1 = substr(ctg_healthy$drug1, 1, 4)
ctg_healthy$drug2 = substr(ctg_healthy$drug2, 1, 4)
heal_com_loewe_diag <- ctg_healthy %>% dplyr::select(sample,drug1,drug2,loewe_diag) %>% group_by(drug1,drug2) %>% summarise(loewe_diag_heal=mean(loewe_diag))
aml_com_loewe_diag <- ctg_aml %>% dplyr::select(sample,drug1,drug2,cluster,group,loewe_diag) %>%
  filter(loewe_diag>0) %>% group_by(drug1,drug2,cluster,group) %>% summarise(loewe_diag_aml=mean(loewe_diag))
ctg_aml_hel <- left_join(aml_com_loewe_diag,heal_com_loewe_diag, by = c("drug1","drug2")) %>% 
 mutate(ratio = loewe_diag_aml/loewe_diag_heal)
#------------------P-value-------------------------------------
data <- ctg_aml_hel %>% filter(loewe_diag_aml > quantile(ctg_aml_hel$loewe_diag_aml,probs =0.25) & loewe_diag_heal<quantile(ctg_aml_hel$loewe_diag_heal,probs =0.75))
prop_test(x = 8, n = 14, p = 0.3, detailed = TRUE,alternative = "greater")
#------------------------------------------------------------
g_loewe_diag <- ggplot(ctg_aml_hel, aes(x = loewe_diag_aml, y = loewe_diag_heal)) +
  geom_point(aes(colour = cluster), size = 5, alpha = 0.6) +
  geom_label_repel(data=filter(ctg_aml_hel, loewe_diag_aml> median(ctg_aml_hel$loewe_diag_aml) & 
                                 loewe_diag_heal < median(ctg_aml_hel$loewe_diag_heal)), 
                   size =4.5, fontface = "bold",alpha = 0.7, 
                   aes(label=paste0(drug1 ,",", drug2)), 
                   nudge_x = .2, 
                   box.padding = 0.5,nudge_y = 1,
                   segment.curvature = -0.1,segment.angle = 20) +
 geom_hline(yintercept = median(ctg_aml_hel$loewe_diag_heal), linetype = "dashed",color = "blue")+
 geom_vline(xintercept = median(ctg_aml_hel$loewe_diag_aml), linetype = "dashed", color = "red")+
 annotate("text", x=7, y=-1, label= "P-value < 0.02", size = 4.5, fontface = "bold") +
 theme(panel.background = element_rect(fill = "white", colour = "grey50"))+
 theme(legend.position = c(0.15, 0.82),legend.title = element_text(size = 8), 
               legend.text = element_text(size = 8))+
 theme(axis.text.y = element_text(face = "bold",size = 14)) +
 theme(axis.text.x = element_text(face = "bold",size = 14)) +
 theme(axis.title.x = element_text(face = "bold",size = 14)) +
 theme(axis.title.y = element_text(face = "bold",size = 14)) +
 scale_color_manual(name = "",labels=c("Cluster 1","Intercluster","Cluster 2"),values = c("#ffcf20FF","#10a53dFF","#541352FF")) +
 theme(legend.text = element_text(size = 8, face = "bold")) +
 xlab("loewe (AML Samples)") + ylab("loewe (Healthy Samples)") 
```

### CR
```{r}
ctg_healthy <- readRDS(file  = "data/rds/ctg_healthy.rds")
ctg_aml <- readRDS(file = "data/rds/ctg_aml.rds")
ctg_aml$drug1 = substr(ctg_aml$drug1, 1, 4)
ctg_aml$drug2 = substr(ctg_aml$drug2, 1, 4)
ctg_healthy$drug1 = substr(ctg_healthy$drug1, 1, 4)
ctg_healthy$drug2 = substr(ctg_healthy$drug2, 1, 4)
heal_com_IR <- ctg_healthy %>% dplyr::select(sample,drug1,drug2,IR) %>% group_by(drug1,drug2) %>% 
  summarise(IR_heal=mean(IR))
aml_com_IR <- ctg_aml %>% dplyr::select(sample,drug1,drug2,cluster,group,IR) %>%
  filter(IR>1.1) %>% group_by(drug1,drug2,cluster,group) %>% summarise(IR_aml=mean(IR))
ctg_aml_hel <- left_join(aml_com_IR,heal_com_IR, by = c("drug1","drug2")) %>% 
  mutate(ratio = IR_aml/IR_heal)
#----------------------------P-value--------------------------------------
data <- ctg_aml_hel %>% filter(IR_aml > quantile(ctg_aml_hel$IR_aml,probs =0.25) & IR_heal<quantile(ctg_aml_hel$IR_heal,probs =0.75))
prop_test(x = 8, n = 13, p = 0.3, detailed = TRUE,alternative = "greater")
#-------------------------------------------------------------------------
g_IR <- ggplot(ctg_aml_hel, aes(x = IR_aml, y = IR_heal)) +
  geom_point(aes(colour = cluster), size = 5, alpha = 0.6) +
  geom_label_repel(data=filter(ctg_aml_hel, IR_aml> median(ctg_aml_hel$IR_aml) & 
                                 IR_heal < median(ctg_aml_hel$IR_heal)), 
                   size =4.5, fontface = "bold",alpha = 0.7, 
                   aes(label=paste0(drug1 ,",", drug2)), 
                   max.overlaps = Inf,nudge_x = 0.1, 
                   box.padding = 0.1,nudge_y = -0.2,
                   segment.curvature = -0.1,segment.angle = 20) +
 geom_hline(yintercept = median(ctg_aml_hel$IR_heal), linetype = "dashed", color = "blue") +
 geom_vline(xintercept = median(ctg_aml_hel$IR_aml), linetype = "dashed", color = "red")+
 annotate("text", x=1.5, y=1, label= "P-value < 0.01", size = 4.5, fontface = "bold") +
 theme(panel.background = element_rect(fill = "white", colour = "grey50"))+
 theme(legend.position = c(0.15, 0.82),legend.title = element_text(size = 8), 
               legend.text = element_text(size = 8))+
 theme(axis.text.y = element_text(face = "bold",size = 14)) +
 theme(axis.text.x = element_text(face = "bold",size = 14)) +
theme(axis.title.x = element_text(face = "bold",size = 14)) +
 theme(axis.title.y = element_text(face = "bold",size = 14)) +
 scale_color_manual(name = "",labels=c("Cluster 1","Intercluster","Cluster 2"),values = c("#ffcf20FF","#10a53dFF","#541352FF")) +
 theme(legend.text = element_text(size = 8, face = "bold")) +
 xlab("CR (AML Samples)") + ylab("CR (Healthy Samples)") 
```

Response and synergy figures
```{r}
library("cowplot")
library("grid")

g <- ggarrange(g_diag, g_zip_hd,g_hsa_hd,g_bliss_hd,g_loewe_diag,g_IR,ncol = 3, nrow = 2,
               labels = c("A","B","C","D","E","F"))
ggsave("data/fig/synergies.pdf",width = 16, height = 10)
```

# Flow-cytometry data analysis

## boxplot of responses on different markers and CTG
```{r}
library(upstartr)
flow_response <- readRDS(file = "data/rds/flow_response.rds")
df <- flow_response %>% 
  filter( ! (marker == "live" | marker == "monocyte-like" | marker== "CD11b+"
              | marker == "singlets" | marker == "cd38+" | 
               (sample == "BERG_47007") & marker == "lymphocytes" ) ) %>% 
  mutate(dc = paste(drug1, "-",drug2)) 

ctg_aml_response <- readRDS(file = "data/rds/ctg_aml_response.rds")
ctg_aml_response$sample[ctg_aml_response$sample == "20220610_7087"] = "FH_7087"
ctg_aml_response$sample[ctg_aml_response$sample == "20220613_47007"] = "BERG_47007"
ctg_aml_response$sample[ctg_aml_response$sample == "20220614_2030"] = "FH_2030"
ctg_aml_response$sample[ctg_aml_response$sample == "20220614_8668"] = "FH_8668"
ctg_aml_response$sample[ctg_aml_response$sample == "20220614_2683"] = "FHRB_2683"
ctg_aml_response = ctg_aml_response[ctg_aml_response$sample %in% flow_response$sample,]
ctg_aml_response = ctg_aml_response %>% mutate(dc = paste(drug1, "-",drug2))
ctg_aml_response = ctg_aml_response[ctg_aml_response$dc %in% df$dc,]
df_blast <- filter(df,marker == "blasts")
ctg_common = merge(ctg_aml_response,df_blast, by.x = c("sample", "dc","dose1"), by.y = c("sample","dc","dose1"))
n= 60
ctg <- data.frame(sample = character(n),
                  marker =rep("CTG",n),
                  drug1 = character(n),
                     drug2=character(n),
                     dose1=double(n),dose2=double(n),response=double(n))
ctg$sample <- ctg_common$sample;
ctg$drug1 <- ctg_common$drug1.x;ctg$drug2 <- ctg_common$drug2.x
ctg$dose1 <- ctg_common$dose1;ctg$dose2 <- ctg_common$dose2.x;
ctg$response <- ctg_common$response.x;
ctg$IR <- ctg_common$IR.x;ctg$zip <- ctg_common$zip.x;
ctg$hsa <- ctg_common$hsa.x;ctg$bliss <- ctg_common$bliss.x;
ctg$loewe <- ctg_common$loewe.x;ctg$dc = ctg_common$dc
ctg_flow <- rbind(df,ctg)
#saveRDS(ctg_flow,file = "data/rds/ctg_flow.rds")
df <- ctg_flow %>% filter(!dose1 ==1)

df <- df %>%
   mutate(dos = case_when(dose1 == 10 ~ 1, 
                          dose1 == 100 ~ 2,
                          dose1 == 1000 ~ 3))

## boxplot and points
gg <- ggplot(df, aes(x=reorder_within(dc, -response,marker), response, color = dc, alpha = dos)) +
    geom_boxplot() +
  scale_x_reordered() +
  facet_wrap(~marker, nrow = 4, scales = "free_x") + 
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(position = position_jitter(width = 0.2, height = 0.2)) + 
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y = element_text(size = 9, face = "bold"),
        legend.position=c(.8, .13)) +
  labs (title = "" ,x= "", y = "Response") +
  scale_color_discrete(name="Drug Combination" ) +
  guides(alpha = FALSE)      

```
## Correlation between CTG and Flowcytometry results

```{r}
library(ggplot2)
library(ggpubr)
library(readxl)
theme_set(theme_pubr())
heatmap_final <- read_excel("data/xls/heatmap.xlsx")
flow_single <- rowMeans(heatmap_final[1:5,3:7], na.rm=T)
flow_comb <- rowMeans(heatmap_final[6:10,3:7], na.rm=T)
ctg_single <- rowMeans(heatmap_final[16:20,3:7], na.rm=T)
ctg_comb <- rowMeans(heatmap_final[21:25,3:7], na.rm=T)
single <- data.frame("CTG" = ctg_single,"FC-BLAST" = flow_single)
combination <- data.frame("CTG" = ctg_comb,"FC-BLAST" = flow_comb)
gs <- ggplot(single, aes(x = ctg_single, y = flow_single)) +
  geom_point(size=3,color ="darkred") +
  theme_bw() +
  theme(axis.text.x = element_text(colour="black", vjust = 0.5, hjust=1,size = 14,face = "bold"),
        axis.text.y = element_text(colour="black",size = 14, face = "bold"),
        axis.title = element_text(size = 14)) +
  labs (title = "Single Drugs" ,x= "CTG", y = "FC-Blast") +
  annotate("text", x = 10, y = 25, label = "cor = 0.97", size =6) +
  geom_smooth(method = lm, se = T)
gc <- ggplot(combination, aes(x = ctg_comb, y = flow_comb)) +
  geom_point(size=3,color ="darkred") +
  theme_bw() +
  theme(axis.text.x = element_text(colour="black", vjust = 0.5, hjust=1,size = 14,face = "bold"),
        axis.text.y = element_text(colour="black",size = 14, face = "bold"),
        axis.title = element_text(size = 14)
        ) +
  labs (title = "Drug Combinations" ,x= "CTG", y = "FC-Blast", sixe =12) +
  annotate("text", x = 18, y = 75, label = "cor = 0.90",size = 6) +
  geom_smooth(method = lm, se = T)
figure <- ggarrange(gs, gc,
                    labels = c("A", "B"),
                    ncol = 1, nrow = 2)

```
